---
title: "Hands-on 2 - DataVis with Advanced ggplot2"
editor: visual
date:  20 Jan 2023
date-modified: "`r Sys.Date()`"
code-copy: true
website: 
    google-analytics: "G-PKMQ2W4ZRC"
format:
  html:
    code-overflow: wrap
    code-fold: true
    code-summary: "Show the code"
    css: styles.css
---

# **1. OVERVIEW**

This study explores data visualisation using [***Tidyverse***](https://www.tidyverse.org/) package.

# **2. R PACKAGE REQUIRED**

## **2.1 Load R Packages**

::: {.callout-warning appearance="simple" icon="false"}
::: {.callout-alert appearance="simple" icon="false"}
[Usage of the code chunk below :]{style="color:#a39f9d"}

[***p_load( )*** - pacman -]{style="color:#d46e15"} to load packages into R environment. This function will attempt to install the package from CRAN or pacman repository list if it is not installed.
:::

```{r}
#| code-fold: false
pacman::p_load(tidyverse, patchwork, ggthemes, hrbrthemes, ggrepel) 
```
:::

# **3. DATA PREPARATION**

## **3.1 Acquire Data Source**

This study will be based on fictitious data set provided with this lesson.

::: {.callout-tip collapse="true" appearance="simple" icon="false"}
## [**Highlights of the data set**]{style="color:#3a9c88"}

1.  Year end examination grades of a cohort of primary 3 students from a local school.

2.  There are a total of seven attributes. Four of them are categorical data type and the other three are in continuous data type.

-   The categorical attributes are: ID, CLASS, GENDER and RACE.
-   The continuous attributes are: MATHS, ENGLISH and SCIENCE.
:::

## **3.2 Import Data**

### **3.2.1 Import Data**

::: {.callout-warning appearance="simple" icon="false"}
Involved two (2) steps : import and inspect imported data set.

::: {.callout-alert appearance="simple" icon="false"}
[Usage of the code chunk below :]{style="color:#a39f9d"}

[***read_csv( )*** - readr -]{style="color:#d46e15"} to read CSV file into a tibble.

[***problems( )*** - readr -]{style="color:#d46e15"} to reveal any parsing errors when importing the CSV file.
:::

```{r}
#| code-fold: false
exam_data <- read_csv("data/Exam_data.csv")

problems(exam_data)
```
:::

### **3.2.2 Inspect Missing Value**

::: {.callout-warning appearance="simple" icon="false"}
```{r}
skimr::skim(exam_data)
```

[Remarks :]{style="color:#d69c3c"}

[No missing data.]{style="color:#d69c3c"}
:::

# **4. DATA VISUALISATION**

## **4.1 Plot with Label & Annotation**

::: {.callout-warning appearance="simple" icon="false"}
text

### **4.1.1 Overlapping Issue**

::: {.callout-alert appearance="simple" icon="false"}
[Usage of the code chunk below :]{style="color:#a39f9d"}

[***xxx( )*** - xxx -]{style="color:#d46e15"} to compute.
:::

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS, 
           y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, 
              size = 0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

### **4.1.2 Solve with ggplot2 :: ggrepel**

::: {.callout-alert appearance="simple" icon="false"}
[Usage of the code chunk below :]{style="color:#a39f9d"}

[***xxx( )*** - xxx -]{style="color:#d46e15"} to compute.
:::

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS, 
           y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, 
              size = 0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

[Remarks :]{style="color:#d69c3c"}

ggrepel: 318 unlabeled data points (too many overlaps). Consider increasing max.overlaps
:::

## **4.2 Theme Options**

This exercise explore 3 themes packages, namely :

1.  ggplot2 package

2.  ggthemes

3.  patchwork package

### **4.2.1 Themes by ggplot2**

::: {.callout-warning appearance="simple" icon="false"}
There are [8 built-in themes for ggplot2](https://ggplot2.tidyverse.org/reference/index.html#themes) :

-   `theme_gray()`,

-   `theme_bw()`,

-   `theme_classic()`,

-   `theme_dark()`,

-   `theme_light()`,

-   `theme_linedraw()`,

-   `theme_minimal(),`

-   `theme_void()`

#### **4.2.1.1 theme_gray( )**

```{r}
#| code-fold: false
ggplot(data = exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins = 20, 
                 boundary = 100,
                 color = "grey25", 
                 fill = "grey90") +
  theme_gray() +  
  ggtitle("Distribution of Maths scores") 
```

#### **4.2.1.2 theme_bw( )**

::: {.callout-alert appearance="simple" icon="false"}
[Usage of the code chunk below :]{style="color:#a39f9d"}

[***xxx( )*** - xxx -]{style="color:#d46e15"} to compute.
:::

```{r}
#| code-fold: false
ggplot(data = exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins = 20, 
                 boundary = 100,
                 color = "grey25", 
                 fill = "grey90") +
  theme_bw() +  
  ggtitle("Distribution of Maths scores") 
```
:::

### **4.2.2 Themes by ggplot2 :: ggthemes**

::: {.callout-warning appearance="simple" icon="false"}
[ggthemes](https://mran.microsoft.com/snapshot/2016-12-28/web/packages/ggthemes/vignettes/ggthemes.html) provides extension to replicate the look of plots by Edward Tufte, Stephen Few, [Fivethirtyeight](https://fivethirtyeight.com/), [The Economist](https://www.economist.com/graphic-detail), 'Stata', 'Excel', and [The Wall Street Journal](https://www.pinterest.com/wsjgraphics/wsj-graphics/), among others.

::: {.callout-alert appearance="simple" icon="false"}
[Usage of the code chunk below :]{style="color:#a39f9d"}

[***xxx( )*** - xxx -]{style="color:#d46e15"} to compute.

::: callout-note
## [**use geom_col( ) if the bars need to represent values in the data.**]{style="color:#3a9c88"}
:::
:::

```{r}
#| code-fold: false
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins = 20, 
                 boundary = 100,
                 color = "grey25", 
                 fill = "grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```

#### 
:::

### **4.2.3 Themes by ggplot2 :: hrbrthemes**

[hrbrthemes](https://cinc.rud.is/web/packages/hrbrthemes/) package provides a base theme that focuses on typographic elements, including where various labels are placed as well as the fonts that are used.

-   The second goal centers around productivity for a production workflow. In fact, this "production workflow" is the context for where the elements of hrbrthemes should be used. Consult [this vignette](https://cran.r-project.org/web/packages/hrbrthemes/vignettes/why_hrbrthemes.html) to learn more.

```{r}
ggplot(data = exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins = 20, 
                 boundary = 100,
                 color = "grey25", 
                 fill = "grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins = 20, 
                 boundary = 100,
                 color = "grey25", 
                 fill = "grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18, 
              base_size = 15, 
              grid = "Y") 
```

**\
What can we learn from the code chunk below?**

-   `axis_title_size` argument is used to increase the font size of the axis title to 18,

-   `base_size` argument is used to increase the default axis label to 15, and

-   `grid` argument is used to remove the x-axis grid lines.

## 

4.3 Composite Plot with Facet Option

Creating Composite Graphics: pathwork methods

It is not unusual that multiple graphs are required to tell a compelling visual story. There are several ggplot2 extensions provide functions to compose figure with multiple graphs. In this section, I am going to shared with you [**patchwork**](https://patchwork.data-imaginist.com/index.html).

Patchwork package has a very simple syntax where we can create layouts super easily. Here's the general syntax that combines: - Two-Column Layout using the Plus Sign +. - Parenthesis () to create a subplot group. - Two-Row Layout using the Division Sign `\`

create composite plot by combining multiple graphs. First, let us create three statistical graphics.\\

```{r}
p1 <- ggplot(data = exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins = 20, 
                 boundary = 100,
                 color = "grey25", 
                 fill = "grey90") + 
  coord_cartesian(xlim = c(0,100)) +
  ggtitle("Distribution of Maths scores")
```

```{r}
p2 <- ggplot(data = exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins = 20, 
                 boundary = 100,
                 color = "grey25", 
                 fill = "grey90") +
  coord_cartesian(xlim = c(0,100)) +
  ggtitle("Distribution of English scores")
```

```{r}
p3 <- ggplot(data = exam_data, 
             aes(x = MATHS, 
                 y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, 
              size = 0.5) +  
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

```{r}
p1 + p2 / p3
```

\| will place the plots beside each other, while / will stack them.

```{r}
(p1 / p2) | p3
```

Plot assembly - https://patchwork.data-imaginist.com/articles/guides/assembly.html

### Working with patchwork

**patchwork** also provides auto-tagging capabilities, in order to identify subplots in text:

```{r}
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```

```{r}
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```

### Working with patchwork

Beside providing functions to place plots next to each other based on the provided layout. With [`inset_element()`](https://patchwork.data-imaginist.com/reference/inset_element.html) of **patchwork**, we can place one or several plots or graphic elements freely on top or below another plot

```{r}
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```

# **7. REFERENCES**

-   T.S. Kam. (2023). Handout lesson material for Hands-on Exercise 2. Singapore Management University. https://isss608-ay2022-23jan.netlify.app/hands-on_ex/hands-on_ex02/hands-on_ex02-beyond_ggplot2#/title-slide

-   [Patchwork R package goes nerd viral](https://www.littlemissdata.com/blog/patchwork)

-   [ggrepel](https://ggrepel.slowkow.com/index.html)

-   [ggthemes](https://ggplot2.tidyverse.org/reference/ggtheme.html)

-   [hrbrthemes](https://cinc.rud.is/web/packages/hrbrthemes/)

-   [ggplot tips: Arranging plots](https://albert-rapp.de/post/2021-10-28-extend-plot-variety/)

-   [ggplot2 Theme Elements Demonstration](https://henrywang.nl/ggplot2-theme-elements-demonstration/)

-   [ggplot2 Theme Elements Reference Sheet](https://isabella-b.com/blog/ggplot2-theme-elements-reference/)
