---
title: "Create Cycle Plot"
subtitle: "Hands-on Exercise 6.2 - Visual Analysis for Time-Oriented Data"
editor: visual
date:  17 Feb 2023
date-modified: "`r Sys.Date()`"
code-copy: true
execute: 
  echo: true
  eval: true
  warning: false
  error: false
website: 
    google-analytics: "G-PKMQ2W4ZRC"
format:
  html:
    code-overflow: wrap
    code-fold: true
    code-summary: "Show the code"
    css: styles.css
---

# **1. OVERVIEW**

This study explores steps to create a cycle plot with tidyverse functions.

# **2. R PACKAGE REQUIRED**

## **2.1 Load R Packages**

```{r}
#| code-fold: false
pacman::p_load(lubridate, scales, viridis, ggthemes, gridExtra, tidyverse, readxl, knitr, data.table)
```

::: {.callout-tip collapse="true" appearance="simple" icon="false"}
## [**R packages highlights :**]{style="color:#3a9c88"}

-   
:::

# **3. DATA PREPARATION**

## **3.1 Acquire Data Source**

This study will be based on *arrivals_by_air.xlsx* data.

## **3.2 Import Data**

### **3.2.1 Import Attribute Data**

Involved two (2) steps : import and inspect imported data set.

::: {.callout-tip collapse="true" appearance="simple" icon="false"}
## [**⇳ Usage of the code chunk below :**]{style="color:#3a9c88"}

[***read_excel( )*** - readr -]{style="color:#d46e15"} to read Excel file into a tibble.

[***problems( )*** - readr -]{style="color:#d46e15"} to reveal any parsing errors when importing the CSV file.
:::

```{r}
air <- read_excel("data/arrivals_by_air.xlsx")

problems(air)
```

### **3.2.2 Examine data**

```{r}
glimpse(air, 70)
```

## 3.3. Wrangle data

### 3.3.1 Derive new fields

Derive "month" and "year" from the "Month-Year" field.

```{r}
air$month <- factor(month
                    (air$`Month-Year`), 
                     levels = 1:12, 
                     labels = month.abb, 
                     ordered = TRUE) 
air$year <- year(ymd(air$`Month-Year`))
```

#### 3.3.1.1 Verify output

```{r}
ls(air)
```

### **3.3.2 Filter data**

Select "Vietnam" data for year 2010 onward.

```{r}
#| code-fold: false
Vietnam <- air %>% 
  select(`Vietnam`, 
         month, 
         year) %>%
  filter(year >= 2010)
```

### **3.3.3 Compute data for geom_hline**

Derive new field name "avgvalue" to reflect the average arrivals for each month.

```{r}
#| code-fold: false
hline.data <- Vietnam %>% 
  group_by(month) %>%
  summarise(avgvalue = mean(Vietnam))
```

#### 3.3.3.1 Inspect output

```{r}
glimpse(hline.data, 70)
```

# 4. CYCLE PLOT

## **4.1 Plot Cycle Plot**

```{r}
#| code-fold: false
#| fig-height: 16
#| fig-width: 12
ggplot() + 
  geom_line(data = Vietnam,
            aes(x = year, 
                y = Vietnam, 
                group = month), 
            colour = "black") +
  geom_hline(aes
             (yintercept = avgvalue), 
             data = hline.data, 
             linetype = 6, 
             colour = "red", 
             size = 0.5) + 
  facet_grid(~month) +
  labs(axis.text.x = element_blank(),
       title = "Visitor arrivals from Vietnam by air, Jan 2010-Dec 2019") +
  xlab("") +
  ylab("No. of Visitors")
```
