---
title: "Ex[4.2] Fundamentals of Visual Analytics - Visualising Models"
editor: visual
date:  3 Feb 2023
date-modified: "`r Sys.Date()`"
code-copy: true
execute: 
  echo: true
  eval: true
  warning: false
website: 
    google-analytics: "G-PKMQ2W4ZRC"
format:
  html:
    code-overflow: wrap
    code-fold: true
    code-summary: "Show the code"
    css: styles.css
---

# **1. OVERVIEW**

This study explores data visualisation using [***Tidyverse***](https://www.tidyverse.org/) package.

In this section, you will learn how to visualise model diagnostic and model parameters by using parameters package.

-   Toyota Corolla case study will be used. The purpose of study is to build a model to discover factors affecting prices of used-cars by taking into consideration a set of explanatory variables.

::: {.callout-tip collapse="true" appearance="simple" icon="false"}
## [**Scope of this study :**]{style="color:#3a9c88"}

In this exercise, learn how to create :

-   ggstatsplot package
:::

# **2. R PACKAGE REQUIRED**

## **2.1 Load R Packages**

```{r}
#| code-fold: false
pacman::p_load(readxl, performance, parameters, see, readr, ggstatsplot) 
```

::: {.callout-tip collapse="true" appearance="simple" icon="false"}
## [**Usage of R Packages :**]{style="color:#3a9c88"}

-   [**ggstatsplot**](https://indrajeetpatil.github.io/ggstatsplot/) is an extension of [**ggplot2**](https://ggplot2.tidyverse.org/) package for creating graphics with details from statistical tests included in the information-rich plots themselves.

    -   To provide alternative statistical inference methods by default.

    -   To follow best practices for statistical reporting. For all statistical tests reported in the plots, the default template abides by the [APA](https://my.ilstu.edu/~jhkahn/apastats.html) gold standard for statistical reporting. For example, here are results from a robust t-test.
:::

# **3. DATA PREPARATION**

## **3.1 Acquire Data Source**

This study will be based on fictitious data set provided with this lesson.

::: {.callout-tip collapse="true" appearance="simple" icon="false"}
## [**Highlights of the data set**]{style="color:#3a9c88"}

1.  x
:::

## **3.2 Import Data**

### **3.2.1 Import Data**

### Importing Excel file: readxl methods

In the code chunk below, [*read_xls()*](https://readxl.tidyverse.org/reference/read_excel.html) of [**readxl**](https://readxl.tidyverse.org/) package is used to import the data worksheet of `ToyotaCorolla.xls` workbook into R.

```{r}
#| code-fold: false
car_resale <- read_xls("data/ToyotaCorolla.xls",
                      "data")

problems(car_resale)
```

Notice that the output object `car_resale` is a tibble data frame.

### Multiple Regression Model using lm()

The code chunk below is used to calibrate a multiple linear regression model by using *lm()* of Base Stats of R.

```{r}
model <- lm(Price ~ Age_08_04 + Mfg_Year + KM + 
              Weight + Guarantee_Period, data = car_resale)
model
```

### Model Diagnostic: checking for multicolinearity:

In the code chunk, [*check_collinearity()*](https://easystats.github.io/performance/reference/check_collinearity.html) of [**performance**](https://easystats.github.io/performance/) package.

```{r}
check_collinearity(model)
```

```{r}
check_c <- check_collinearity(model)
plot(check_c)
```

\

### Model Diagnostic: checking normality assumption

In the code chunk, [*check_normality()*](https://easystats.github.io/performance/reference/check_normality.html) of [**performance**](https://easystats.github.io/performance/) package.

```{r}
model1 <- lm(Price ~ Age_08_04 + KM + 
              Weight + Guarantee_Period, data = car_resale)
```

```{r}
check_n <- check_normality(model1)
```

```{r}
plot(check_n)
```

### Model Diagnostic: Check model for homogeneity of variances

In the code chunk, [*check_heteroscedasticity()*](https://easystats.github.io/performance/reference/check_heteroscedasticity.html) of [**performance**](https://easystats.github.io/performance/) package.

```{r}
check_h <- check_heteroscedasticity(model1)
```

```{r}
plot(check_h)
```

### Model Diagnostic: Complete check

We can also perform the complete by using [*check_model()*](https://easystats.github.io/performance/reference/check_model.html).

```{r}
check_model(model1)
```

### Visualising Regression Parameters: see methods

In the code below, plot() of see package and parameters() of parameters package is used to visualise the parameters of a regression model.

```{r}
plot(parameters(model1))
```

### Visualising Regression Parameters: *ggcoefstats()* methods

In the code below, [*ggcoefstats()*](https://indrajeetpatil.github.io/ggstatsplot/reference/ggcoefstats.html) of ggstatsplot package to visualise the parameters of a regression model.

```{r}
ggcoefstats(model1, 
            output = "plot")
```
