---
title: "In-Class Ex02"
editor: visual
date:  21 Jan 2023
date-modified: "`r Sys.Date()`"
code-copy: true
website: 
    google-analytics: "G-PKMQ2W4ZRC"
format:
  html:
    code-overflow: wrap
    code-fold: true
    code-summary: "Show the code"
    css: styles.css
---

# **1. OVERVIEW**

This study explores data visualisation using [***Tidyverse***](https://www.tidyverse.org/) package.

# **2. R PACKAGE REQUIRED**

## **2.1 Load R Packages**

::: {.callout-warning appearance="simple" icon="false"}
::: {.callout-alert appearance="simple" icon="false"}
[Usage of the code chunk below :]{style="color:#a39f9d"}

[***p_load( )*** - pacman -]{style="color:#d46e15"} to load packages into R environment. This function will attempt to install the package from CRAN or pacman repository list if it is not installed.
:::

```{r}
#| code-fold: false
pacman::p_load(tidyverse, patchwork, ggthemes, hrbrthemes, ggrepel) 
```
:::

# **3. DATA PREPARATION**

## **3.1 Acquire Data Source**

This study will be based on fictitious data set provided with this lesson.

::: {.callout-tip collapse="true" appearance="simple" icon="false"}
## [**Highlights of the data set**]{style="color:#3a9c88"}

1.  Year end examination grades of a cohort of primary 3 students from a local school.

2.  There are a total of seven attributes. Four of them are categorical data type and the other three are in continuous data type.

-   The categorical attributes are: ID, CLASS, GENDER and RACE.
-   The continuous attributes are: MATHS, ENGLISH and SCIENCE.
:::

## **3.2 Import Data**

### **3.2.1 Import Data**

::: {.callout-warning appearance="simple" icon="false"}
Involved two (2) steps : import and inspect imported data set.

```{r}
#| code-fold: false
exam_data <- read_csv("data/Exam_data.csv")

problems(exam_data)
```
:::

### 
