---
title: "BSW : Building Better Explanatory Models"
editor: visual
date:  18 Jan 2023
date-modified: "`r Sys.Date()`"
code-copy: true
website: 
    google-analytics: "G-PKMQ2W4ZRC"
format:
  html:
    code-overflow: wrap
    code-fold: true
    code-summary: "Show the code"
    css: styles.css
---

# **1. OVERVIEW**

An overview of explanatory model

# 2. R PACKAGE REQUIRED

## 2.1 Load R Packages

```{r}
pacman::p_load(tidyverse, readxl, report, performance, funModeling, parameters, gtsummary)
```

# **3. DATA**

## **3.1 Acquire Data Source**

This study will be based on fictitious data set provided with this lesson.

::: {.callout-tip collapse="true" appearance="simple" icon="false"}
## [**Highlights of the data set**]{style="color:#3a9c88"}

1.  x
:::

## **3.2 Import Data**

### **3.2.1 Read Data into R**

```{r}
#| code-fold: false
car_resale <- read_xls("data/ToyotaCorolla.xls",
                       "data")

problems(car_resale)
```

### **3.2.2 Inspect Data**

#### **3.2.2.1 identify variable type**

```{r}
glimpse(car_resale,60)
```

#### **3.2.2.2 check for missing data**

```{r}
skimr::skim(car_resale)
```

#### **3.2.2.3 reveal unique value**

When use funModeling function, turn the "tbl_df" into "data.frame".

```{r}
#status <- df_status(car_resale) << alternate way to code

car_resale %>% 
  df_status()
```

## 3.3 Data Wrangling

### 3.3.1 Convert Variable Type

```{r}
cols <- c("Mfg_Month", "HP_Bin", "CC_bin", "Doors", "Gears", "Cylinders", "Fuel_Type", "Color", "Met_Color", "Automatic", "Mfr_Guarantee", "BOVAG_Guarantee", "ABS", "Airbag_1", "Airbag_2", "Airco", "Automatic_airco", "Boardcomputer", "CD_Player", "Central_Lock", "Powered_Windows", "Power_Steering", "Radio", "Mistlamps", "Sport_Model", "Backseat_Divider", "Metallic_Rim", "Radio_cassette", "Tow_Bar")


car_resale <- read_xls("data/ToyotaCorolla.xls", 
                       sheet = "data") %>%
  mutate(Id = as.character(Id)) %>%
  mutate_each_(funs(factor(.)),
               cols)
```

#### 3.3.1.1 inspect output

```{r}
profiling_num(car_resale)
```

#### 3.3.1.2 EDA for Continuous Variables

```{r}
plot_num(car_resale)
```

#### 3.3.1.3 EDA for Character Variables

```{r}
freq(car_resale[3:38])
```

# 4. MODEL DIAGNOSTIC

## 4.1 Create Model

```{r}
model <- lm(Price ~ Age_08_04 + 
              Mfg_Year + 
              KM + 
              Weight + 
              Guarantee_Period, 
            data = car_resale)

model
```

## 4.2 Diagnose Model's Multicollinearity

### 4.2.1 Check Multicollinearity

```{r}
check_collinearity(model)
```

https://easystats.github.io/easystats/

### 4.2.2 EDA for Multicollinearity

```{r}
check_c <- check_collinearity(model)

plot(check_c)
```

## 4.3 Normality Analysis

### 4.3.1 Create Model for Normality Analysis

```{r}
model1 <- lm(Price ~ Age_08_04 +
               KM + 
               Weight + 
               Guarantee_Period, 
             data = car_resale)

check_normality(model1)
```

### 4.3.2 EDA Model

```{r}
check_n <- check_normality(model1) 
plot(check_n)
```

## 4.4 Model Inspection on Homogeneity of Variances

### 4.4.1 Overview

```{r}
check_h <- check_heteroscedasticity(model1)
plot(check_h)
```

### 4.4.2 Complete Inspection

```{r}
check_model(model1)
```

# 5. SUMMARY

## 5.1 Summarise with summary( ) 

```{r}
summary(model1)
```

## 5.2 Summaries with gtsummary:: tbl_regression() 

### 5.2.1 Summaries Model1

```{r}
tbl_regression(model1,
               intercept = TRUE)
```

### 5.2.2 Include Additional Parameters

```{r}
tbl_regression(model1,
               intercept = TRUE) %>%
  add_glance_source_note(
    label = list(sigma ~ "\U03C3"),
    include = c(r.squared, 
                adj.r.squared,
                AIC, 
                statistic,
                p.value, 
                sigma))
```

## 5.3 Create Report for Model1

```{r}
#| echo: false
#| eval: false
report(model1)
```

We fitted a linear model (estimated using OLS) to predict Price with Age_08_04 (formula: Price \~

Age_08_04 + KM + Weight + Guarantee_Period). The model explains a statistically significant and

substantial proportion of variance (R2 = 0.85, F(4, 1431) = 2004.93, p \< .001, adj. R2 = 0.85). The

model's intercept, corresponding to Age_08_04 = 0, is at -2185.52 (95% CI \[-4092.59, -278.45\], t(1431)

= -2.25, p = 0.025). Within this model:

 - The effect of Age 08 04 is statistically significant and negative (beta = -119.49, 95% CI \[-124.91,

-114.08\], t(1431) = -43.29, p \< .001; Std. beta = -0.61, 95% CI \[-0.64, -0.59\])

 - The effect of KM is statistically significant and negative (beta = -0.02, 95% CI \[-0.03, -0.02\],

t(1431) = -20.04, p \< .001; Std. beta = -0.25, 95% CI \[-0.27, -0.22\])

 - The effect of Weight is statistically significant and positive (beta = 19.72, 95% CI \[18.08, 21.36\],

t(1431) = 23.53, p \< .001; Std. beta = 0.29, 95% CI \[0.26, 0.31\])

 - The effect of Guarantee Period is statistically significant and positive (beta = 26.82, 95% CI

\[2.08, 51.56\], t(1431) = 2.13, p = 0.034; Std. beta = 0.02, 95% CI \[1.73e-03, 0.04\])

Standardized parameters were obtained by fitting the model on a standardized version of the dataset.

95% Confidence Intervals (CIs) and p-values were computed using a Wald t-distribution approximation.,

We fitted a linear model (estimated using OLS) to predict Price with KM (formula: Price \~ Age_08_04 +

KM + Weight + Guarantee_Period). The model explains a statistically significant and substantial

proportion of variance (R2 = 0.85, F(4, 1431) = 2004.93, p \< .001, adj. R2 = 0.85). The model's

intercept, corresponding to KM = 0, is at -2185.52 (95% CI \[-4092.59, -278.45\], t(1431) = -2.25, p =

0.025). Within this model:

 - The effect of Age 08 04 is statistically significant and negative (beta = -119.49, 95% CI \[-124.91,

-114.08\], t(1431) = -43.29, p \< .001; Std. beta = -0.61, 95% CI \[-0.64, -0.59\])

 - The effect of KM is statistically significant and negative (beta = -0.02, 95% CI \[-0.03, -0.02\],

t(1431) = -20.04, p \< .001; Std. beta = -0.25, 95% CI \[-0.27, -0.22\])

 - The effect of Weight is statistically significant and positive (beta = 19.72, 95% CI \[18.08, 21.36\],

t(1431) = 23.53, p \< .001; Std. beta = 0.29, 95% CI \[0.26, 0.31\])

 - The effect of Guarantee Period is statistically significant and positive (beta = 26.82, 95% CI

\[2.08, 51.56\], t(1431) = 2.13, p = 0.034; Std. beta = 0.02, 95% CI \[1.73e-03, 0.04\])

Standardized parameters were obtained by fitting the model on a standardized version of the dataset.

95% Confidence Intervals (CIs) and p-values were computed using a Wald t-distribution approximation.,

We fitted a linear model (estimated using OLS) to predict Price with Weight (formula: Price \~

Age_08_04 + KM + Weight + Guarantee_Period). The model explains a statistically significant and

substantial proportion of variance (R2 = 0.85, F(4, 1431) = 2004.93, p \< .001, adj. R2 = 0.85). The

model's intercept, corresponding to Weight = 0, is at -2185.52 (95% CI \[-4092.59, -278.45\], t(1431) =

-2.25, p = 0.025). Within this model:

 - The effect of Age 08 04 is statistically significant and negative (beta = -119.49, 95% CI \[-124.91,

-114.08\], t(1431) = -43.29, p \< .001; Std. beta = -0.61, 95% CI \[-0.64, -0.59\])

 - The effect of KM is statistically significant and negative (beta = -0.02, 95% CI \[-0.03, -0.02\],

t(1431) = -20.04, p \< .001; Std. beta = -0.25, 95% CI \[-0.27, -0.22\])

 - The effect of Weight is statistically significant and positive (beta = 19.72, 95% CI \[18.08, 21.36\],

t(1431) = 23.53, p \< .001; Std. beta = 0.29, 95% CI \[0.26, 0.31\])

 - The effect of Guarantee Period is statistically significant and positive (beta = 26.82, 95% CI

\[2.08, 51.56\], t(1431) = 2.13, p = 0.034; Std. beta = 0.02, 95% CI \[1.73e-03, 0.04\])

Standardized parameters were obtained by fitting the model on a standardized version of the dataset.

95% Confidence Intervals (CIs) and p-values were computed using a Wald t-distribution approximation.

and We fitted a linear model (estimated using OLS) to predict Price with Guarantee_Period (formula:

Price \~ Age_08_04 + KM + Weight + Guarantee_Period). The model explains a statistically significant

and substantial proportion of variance (R2 = 0.85, F(4, 1431) = 2004.93, p \< .001, adj. R2 = 0.85).

The model's intercept, corresponding to Guarantee_Period = 0, is at -2185.52 (95% CI \[-4092.59,

-278.45\], t(1431) = -2.25, p = 0.025). Within this model:

 - The effect of Age 08 04 is statistically significant and negative (beta = -119.49, 95% CI \[-124.91,

-114.08\], t(1431) = -43.29, p \< .001; Std. beta = -0.61, 95% CI \[-0.64, -0.59\])

 - The effect of KM is statistically significant and negative (beta = -0.02, 95% CI \[-0.03, -0.02\],

t(1431) = -20.04, p \< .001; Std. beta = -0.25, 95% CI \[-0.27, -0.22\])

 - The effect of Weight is statistically significant and positive (beta = 19.72, 95% CI \[18.08, 21.36\],

t(1431) = 23.53, p \< .001; Std. beta = 0.29, 95% CI \[0.26, 0.31\])

 - The effect of Guarantee Period is statistically significant and positive (beta = 26.82, 95% CI

\[2.08, 51.56\], t(1431) = 2.13, p = 0.034; Std. beta = 0.02, 95% CI \[1.73e-03, 0.04\])

Standardized parameters were obtained by fitting the model on a standardized version of the dataset.

95% Confidence Intervals (CIs) and p-values were computed using a Wald t-distribution approximation.
