---
title: "Exploration 2 - Pyramid Chart"
editor: visual
date:  28 Dec 2022
code-copy: true
website: 
    google-analytics: "G-PKMQ2W4ZRC"
format:
  html:
    code-overflow: wrap
    code-fold: true
    code-summary: "Show the code"
    css: styles.css
---

# 1. OVERVIEW

This study explores the **Pyramid chart** using **ggplot2** and supporting R packages.

::: {.callout-tip collapse="true" appearance="simple" icon="false"}
## [**Challenges in wrangling the raw data :**]{style="color:#3a9c88"}

The age group variable, "AG", is a character data type.

If it is used to plot the population pyramid, "10_to_14" will be parsed first instead of "5_to_9" after "0_to_4 is parsed.

Because of that, the order in the chart will be incorrect.

[Resolve manually with these functions - *mutate( )* and *arrange( )*.]{style="color:#d69c3c"}
:::

# 2. R PACKAGE REQUIRED

The following are the packages required for this exercise :

## 2.1 Load R Packages

::: {.callout-warning appearance="simple" icon="false"}
[Usage of the code chunk below :]{style="color:#a39f9d"}

[***p_load( )*** - pacman -]{style="color:#d46e15"} to load packages into R environment. This function will attempt to install the package from CRAN or pacman repository list if it is not installed.

```{r}
pacman::p_load(tidyverse, readxl, knitr, plotly, skimr, questionr, funModeling)
```

[Remarks :]{style="color:#d69c3c"}

[sf, tidyverse, questionr, janitor, psych, ggplot2, gcookbook, tmap, ggpubr, egg, corrplot, gtsummary, regclass, caret, heatmaply, ggdendro, cluster, factoextra, spdep, ClustGeo, GGally, skimr, stringr, funModeling, knitr, caTools, viridis, rgeoda, cowplot, patchwork.]{style="color:#d69c3c"}
:::

# 3. DATA

## 3.1 Acquire Data Source

-   Singstat Dataset for Population

    1.  Download "Singapore Residents by Planning Area Subzone, Age Group, Sex and Type of Dwelling, June 2022" [dataset](https://www.singstat.gov.sg/find-data/search-by-theme/population/geographic-distribution/latest-data) in csv format from Singstat.gov.sg.

## 3.2 Import Data

### 3.2.1 Import Tableau Data

::: {.callout-warning .column-page-right appearance="simple" icon="false"}
Two (2) steps : import and inspect imported data set.

#### 3.2.1.1 import data

::: {.callout-alert appearance="simple" icon="false"}
[Usage of the code chunk below :]{style="color:#a39f9d"}

[***read_csv( )*** - readr -]{style="color:#d46e15"} to read CSV file into a tibble.

[***problems( )*** - readr -]{style="color:#d46e15"} to reveal any parsing errors when importing the CSV file.
:::

```{r}
pop_sgp <- read_csv("data/respopagesexfa2022.csv")

problems(pop_sgp)
```

#### 3.2.1.2 inspect missing value

```{r}
skim(pop_sgp)
```

[Remarks :]{style="color:#d69c3c"}

[Preliminary data interpretation - based on 2022 consensus data, .]{style="color:#d69c3c"}
:::

## 3.3 Data Wrangling

### 3.3.1 Compute Frequency Count

::: {.callout-warning .column-page-right appearance="simple" icon="false"}
Compute the frequency count by "age" and "sex".

#### **3.3.1.1 compute returns by "Sub-category"**

::: {.callout-alert appearance="simple" icon="false"}
[Usage of the code chunk below :]{style="color:#a39f9d"}

[***group_by( )*** - dplyr -]{style="color:#d46e15"} to group Singapore population by age and gender.

[***summarise( )*** - dplyr -]{style="color:#d46e15"} to count the number of residents for each group.

[***arrange( )*** - dplyr -]{style="color:#d46e15"} to sort the rows in descending order of the counted "Returned" value of each "Sub-category".
:::

```{r}
order <- c("0_to_4", "5_to_9", "10_to_14", "15_to_19", "20_to_24", "25_to_29", "30_to_34", "35_to_39", "40_to_44", "45_to_49", "50_to_54", "55_to_59", "60_to_64", "65_to_69", "70_to_74", "75_to_79", "80_to_84", "85_to_89", "90_and_over")

pop_agSx <- pop_sgp %>%
  group_by(`AG`,`Sex`) %>%
  summarise(`Count` = sum(`Pop`)) %>%
  mutate(AG = factor(AG, levels = order)) %>%
  arrange(AG) %>%
  ungroup()

head(pop_agSx)
```
:::

# 4 EXPLORATORY DATA ANALYSIS (EDA)

## 4.1 Plot Pyramid Chart

#### **4.1.1 Plot Static Pyramid**

::: {.callout-warning .column-page-right appearance="simple" icon="false"}
::: {.callout-alert appearance="simple" icon="false"}
[Usage of the code chunk below :]{style="color:#a39f9d"}

[***geom_col( )*** - ggplot2 -]{style="color:#d46e15"} to create the bar chart with the bar length represent the actual population stats.

[***scale_x\_continuous( )*** - ggplot2 -]{style="color:#d46e15"} to modify the labels on the x-axis to provide better readability.

[***ifelse( )*** - base -]{style="color:#d46e15"} to convert population count of Male residents to be negative, so that their data will be plotted on the left side of the pyramid.
:::

```{r}
pyramid_agSx <- ggplot(pop_agSx, 
                       aes(x = ifelse(Sex == "Males",
                                      yes = -Count,
                                      no = Count),
                           y = AG,
                           fill = Sex)) + 
  geom_col() +
  scale_x_continuous(breaks = seq(-150000, 150000, 50000), 
                     labels = paste0(
                       as.character(
                         c(seq(150, 0, -50), 
                           seq(50, 150, 50))),
                       "k")) +
  labs (x = "Population", 
        y = "Age", 
        title = 'Singapore Age-Sex Population Pyramid 2022') +
  theme_bw() +
  theme(axis.ticks.y = element_blank()) +
  scale_fill_manual(values = c("Males" = "lightblue", 
                               "Females" = "lightpink")) 

pyramid_agSx
```
:::

#### **4.1.2 Interactive Pyramid Chart**

::: {.callout-warning .column-page-right appearance="simple" icon="false"}
```{r}
ggplotly(pyramid_agSx, 
         session = "knitr")
```

[Remark :]{style="color:#d69c3c"}

[The pyramid above shows the younger generation's lower percentage, making Singapore's population structure a constrictive type. That means an ageing population.]{style="color:#d69c3c"}

[By June 2022, 4.07 million residents population with a median age of 42.1,]{style="color:#d69c3c"}[^1] [which is approximately a 19% increase from the median age of 35.3 back in the year 2002]{style="color:#d69c3c"}[^2].

[Also, another observation is from 25 age onward, the female population is higher than the male population. As of 2020, Singapore's adjusted gender pay gap (GPG) is at 6%]{style="color:#d69c3c"}[^3].

[Measures and new schemes may be required to further narrow the gender pay gap and encourage the birth rate or may have long-lasting socio-economic repercussions.]{style="color:#d69c3c"}
:::

[^1]: [Singstat. (2022). Population Dashboard. https://www.singstat.gov.sg/find-data/search-by-theme/population/population-and-population-structure/visualising-data/population-dashboard]{style="color:#d69c3c"}

[^2]: [Singstat. (2002). Indicators On Population. https://tablebuilder.singstat.gov.sg/table/TS/M810001]{style="color:#d69c3c"}

[^3]: [Ministry of Manpower. (2020). Report: Singapore's Adjusted Gender Pay Gap. https://stats.mom.gov.sg/Pages/Singapores-Adjusted-Gender-Pay-Gap.aspx]{style="color:#d69c3c"}

# 5. REFERENCES
